Demo: Counterpoint by Construction

### 2019-05-19 doc/farm/farm-abstract.tex
rough draft

%% Title information
\title{Musical Ornaments}
\subtitle{Functional Pearl}

@@

%% Author with single affiliation.
\author{John Leo}
\affiliation{
  \institution{Halfaya Research}            %% \institution is required
  \city{Bellevue}
  \state{WA}
  \country{USA}
}
\email{leo@halfaya.org}          %% \email is recommended

\author{Youyou Cong}
\affiliation{
  \institution{Tokyo Institute of Technology}            %% \institution is required
  \city{Tokyo}
  \country{Japan}
}
\email{cong@c.titech.ac.jp}          %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

@@

\section{Introduction}

Text of paper \ldots

%% Acknowledgments
\begin{acks}                            %% contents suppressed with 'anonymous'
\end{acks}

%% Bibliography
%\bibliography{bibfile}

>>>>

%% Title information
\title{Demo: Counterpoint by Construction}
%\subtitle{Functional Pearl}

@@

%% Author with single affiliation.
\author{Youyou Cong}
\affiliation{
  \institution{Tokyo Institute of Technology}            %% \institution is required
  \city{Tokyo}
  \country{Japan}
}
\email{cong@c.titech.ac.jp}          %% \email is recommended

\author{John Leo}
\affiliation{
  \institution{Halfaya Research}            %% \institution is required
  \city{Bellevue}
  \state{WA}
  \country{USA}
}
\email{leo@halfaya.org}          %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Western music of the common practice period tends to loosely follow sets of rules which
were developed over time to help ensure the aethetic quality of the composition. Rules in
developed in particular for harmony (\cite{piston1987harmony}) and counterpoint (\cite{fux1965study}),
and continue to be taught to music students
both to to understand the music of that time and because they continue to provide a foundation for
modern art and popular music. To help both analyze and synthesize
tonal music, it is worthwhile to attempt to encode these rules into a programming
language, with functional programming languages being particularly wells-suited to this task.
Recently quite a bit of work has been done in Haskell (\cite{fmmh, hihseufha, faamh, hafha, fghm})
to encode the rules of harmony, but little work seems to have been done with counterpoint (\cite{mezzo}).
Furthermore Haskell's type system, lacking full dependent types, is not strong enough to encode
the required logic in a natural way, and requires using a number of extensions.

In this demonstration of work in progress, we present Music Tools (\cite{MusicTools}), a library of
small tools that can be combined functionally to help analyze and synthesize music. The library is
written in the dependently typed language Agda (with Haskell used for I/O), which allows a simple
and natural encoding of rules. We demonstrate its use in writing species counterpoint following
Fux (\cite{fux1965study}), in which type system ensures by construction that the rules are followed.
We show how this both aids human composition and allows for computer-generated creation of correct
counterpoint. We also contrast this work to recent use of machine learnring to generated natural-sounding
counterpoint which does not necessarily follow specific rules (\cite{CounterpointByConvolution}). Finally
we indicate future planned work including extension to handle functional harmony.
\end{abstract}

@@

%% Acknowledgments
\begin{acks}                            %% contents suppressed with 'anonymous'
  Thanks to the participants of the Tokyo Agda Implmentors' Meeting, especially Ulf Norell and Jesper Cockx,
  for many helpful suggestions that improved our Agda code.
\end{acks}

%% Bibliography
\bibliography{farm-abstract.bib}
